RDF Validation of Metadata \\ on Person-Level and Aggregated Data

### 2015-03-03 paper/paper7/paper7.tex
4.2

\textbf{Variables.}
Each variable should have a variable representation (\emph{R-86}; $\mathcal{SL}_{1}$), which is either an (un)ordered code list or a union of XML Schema datatypes.
In case of a code list, associated categories should be stated (\emph{R-71}; $\mathcal{SL}_{0}$).
As the variable \emph{Education at pre-school} has an associated code list, the constraint is not violated.
If a {\em skos:Concept} represents a code (having a {\em skos:notation} property) and a category (having a {\em skos:prefLabel} property), 
then the property {\em disco:isValid} has to be stated indicating if the code is valid (\emph{true}) or missing (\emph{false}) (\emph{R-71}; $\mathcal{SL}_{1}$).

\textbf{Theoretical Concepts.}
Variables must have at least one relationship to a theoretical concept ({\emph{R-75: minimum qualified cardinality restrictions}).
%The variables \emph{number of school years} and \emph{highest educational degree}, e.g., are associated with the theoretical concept \emph{education}. 
The variable \emph{Education at pre-school} is associated with the theoretical concept \emph{education}. 
The default severity level of this constraint is \emph{warning} and not \emph{error}, as in most cases research can be continued without an associated theoretical concept of a given variable.

\textbf{Comparison.}
A very common research question is to compare variables (\emph{comparison}) of multiple studies or countries.
To compare variables, 
(1) variables and (2) variable definitions must be present,
(3) code lists must be structured properly,
(4) for each code an associated category (human-readable label) must be specified, and
(5) code lists must either be identical or at least similar.
If a researcher wants to get a first overview over comparable variables (use case 1), 
the first three constraints may be sufficient for this purpose.
Thus, the severity level of the first three constraints is stronger than the severity level of the next constraints.
If the intention of the researcher is to perform more detailed comparisons (use case 2), however, the violation of the remaining constraints is also more serious.

\textbf{Reusability.}
Constraints of the constraint type \emph{Context-Specific Exclusive OR of Property Groups} (\emph{R-11})
restrict individuals of given classes to have only one of multiple property groups.
Within the context of \emph{Disco}, \emph{skos:Concept}s can have either \emph{skos:definition} (when interpreted as theoretical concepts) or \emph{skos:notation} and \emph{skos:prefLabel} properties (when interpreted as codes and categories), but not both.

\begin{DL}
Concept $\sqsubseteq$ ($\neg$ D $\sqcap$ C) $\sqcup$ (D $\sqcap$ $\neg$ C) \\ 
D $\equiv$ A $\sqcap$ B \\
A $\sqsubseteq$ $\geq$ 1 notation.string $\sqcap$ $\leq$ 1 notation.string \\
B $\sqsubseteq$ $\geq$ 1 prefLabel.string $\sqcap$ $\leq$ 1 prefLabel.string \\
C $\sqsubseteq$ $\geq$ 1 definition.string $\sqcap$ $\leq$ 1 definition.string \\
\end{DL}

@@

\begin{DL}
SummaryStatistics $\sqsubseteq$ $\forall summaryStatisticType.A$ \\
$A \equiv Concept \sqcap \forall inScheme . B$ \\
$B \equiv ConceptScheme \sqcap \{SummaryStatisticType\}$
\end{DL}

@@

@@

In \emph{Missy}, we provide comprehensive Linked Data services\footnote{\textcolor{red}{\url{www.gesis.org/en/missy/linked-data}}} like a \textcolor{red}{triple store} and diverse RDF exports of person-level metadata conforming to the \emph{Disco} vocabulary in multiple concrete syntaxes (e.g. turtle).

>>>>

\textbf{Variables.}
Each variable should have a variable representation (\emph{R-86}; $\mathcal{SL}_{1}$), which is either an (un)ordered code list or a union of datatypes.
In case of a code list, associated categories should be stated (\emph{R-71}; $\mathcal{SL}_{0}$).
The variable \emph{Education at pre-school} has an associated code list, but without any associated categories.
If a {\em skos:Concept} represents a code (having a {\em skos:notation} property) and a category (having a {\em skos:prefLabel} property), 
then the property {\em disco:isValid} has to be stated indicating if the code is valid (\emph{true}) or missing (\emph{false}) (\emph{R-71}; $\mathcal{SL}_{1}$).
Variables must have at least one relationship to a theoretical concept ({\emph{R-75: minimum qualified cardinality restrictions}; $\mathcal{SL}_{1}$).
%The variables \emph{number of school years} and \emph{highest educational degree}, e.g., are associated with the theoretical concept \emph{education}. 
The variable \emph{Education at pre-school} is associated with the theoretical concept \emph{Child Care}. 
The default severity level of this constraint is $\mathcal{SL}_{1}$ and not $\mathcal{SL}_{2}$, as in most cases research can be continued without associated theoretical concepts of given variables.

\textbf{Variable Comparison.}
A very common research question is to compare variables of multiple studies or countries (\emph{comparison}).
To compare variables, 
(1) variables and (2) variable definitions must be present,
(3) code lists must be structured properly,
(4) for each code an associated category (human-readable label) must be specified, and
(5) code lists must either be identical or at least similar.
If a researcher wants to get a first overview over comparable variables (use case 1), 
the first three constraints may be sufficient for this purpose.
Thus, the severity level of the first three constraints is stronger ($\mathcal{SL}_{2}$) than the severity level of the next two constraints ($\mathcal{SL}_{1}$).
If the intention of the researcher is to perform more detailed comparisons (use case 2), however, the violation of the remaining constraints is getting more serious ($\mathcal{SL}_{2}$).

@@

\begin{center}
\begin{DL}
SummaryStatistics $\sqsubseteq$ $\forall summaryStatisticType.A$ \\
$A \equiv Concept \sqcap \forall inScheme . B$ \\
$B \equiv ConceptScheme \sqcap \{SummaryStatisticType\}$
\end{DL}
\end{center}

@@

\textbf{Reusability.}
Within the context of \emph{Disco}, \emph{skos:Concept}s can have either \emph{skos:definition} (when interpreted as theoretical concepts) or \emph{skos:notation} and \emph{skos:prefLabel} properties (when interpreted as codes and categories), but not both ($\mathcal{SL}_{2}$).
Constraints of the constraint type \emph{context-specific exclusive or of property groups} (\emph{R-11})
restrict individuals of given classes to have exactly one of multiple property groups.

\begin{center}
\begin{DL}
Concept $\sqsubseteq$ ($\neg$ D $\sqcap$ C) $\sqcup$ (D $\sqcap$ $\neg$ C), D $\equiv$ A $\sqcap$ B \\
A $\sqsubseteq$ $\geq$ 1 notation.string $\sqcap$ $\leq$ 1 notation.string \\
B $\sqsubseteq$ $\geq$ 1 prefLabel.string $\sqcap$ $\leq$ 1 prefLabel.string \\
C $\sqsubseteq$ $\geq$ 1 definition.string $\sqcap$ $\leq$ 1 definition.string \\
\end{DL}
\end{center}

@@

In \emph{Missy}, we provide comprehensive Linked Data services\footnote{\textcolor{red}{\url{http://www.gesis.org/missy/eu/general-information/linked-data/}}} like a \textcolor{red}{triple store} and diverse RDF exports of person-level metadata conforming to the \emph{Disco} vocabulary in multiple concrete syntaxes (e.g. turtle).
